---
title: "project"
author: "LUQIN GAN"
date: "12/3/2018"
output: html_document
---

## Data exploration

```{r}

library(ggplot2)
library(tidyverse)
library(reshape)
library(knitr)
library(kableExtra)
library(plotly)
train = read.csv('/Users/alice/Documents/yr2term1/data science/project/Data/train_US_1year_nojson.csv')
```

#### Check features with missing values 
```{r}
train$totalTransactionRevenue[which(is.na(train$totalTransactionRevenue))] = 0
miss = data.frame(Features = colnames(train), Number = sapply(1:ncol(train), function(x) length(which(is.na(train[,x])))), Percent = round(sapply(1:ncol(train), function(x) length(which(is.na(train[,x]))))/nrow(train),2))

miss %>%
  mutate_if(is.numeric, function(x) {
    cell_spec(x, bold = T, 
              color = spec_color(x, end = 0.9),
              font_size = spec_font_size(x))
  }) %>%
  kable(escape = F, align = "c",caption = "Missing data summary") %>%
  kable_styling(c("striped", "condensed"), full_width = F)

dat = train[,which(miss$Percent<0.8)]
```

We droped the features which have greater than 80% missing values. 

#### Channel Grouping 

Here we explore the marketing channel grouping distribution. 

```{r}

cg = data.frame(table(train$channelGrouping))
percent <- function(x, digits = 2, format = "f", ...) {
  paste0(formatC(100 * x, format = format, digits = digits, ...), "%")
}
cg  = cg%>% 
  mutate(percent = percent(Freq/sum(Freq)))

# ggplot(cg,aes(x = reorder(Var1,-Freq) , y =Freq))+
#   geom_bar(stat="identity",fill="steelblue")+
#       geom_text(aes(label=percent ), vjust=0) +
#   theme(plot.title = element_text(hjust = 0.5),
#     axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
#     labs(title = "Channel Grouping", x = 'Channel Grouping', y = "Number")
plot_ly(cg,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = cg$percent)%>%
  layout(title = "Channel Grouping",
         xaxis = list(title = "Channel Grouping"),
        yaxis = list(title = "Number"))
```
Over 40% of the visitors used organic search. 

Here we explore the browser distribution. 
```{r}
train_brow = train[which(train$browser%in%names(sort(table(train$browser),decreasing=TRUE)[1:10])),]
brow = data.frame(table(train_brow$browser))
brow = brow[-which(brow$Freq==0),]
brow  = brow%>% 
  mutate(percent = percent(Freq/sum(Freq)))

ggplot(brow,aes(x = reorder(Var1,-Freq) , y =Freq))+
  geom_bar(stat="identity",fill="#FF6666")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
        geom_text(aes(label=percent ), vjust=0) +
    labs(title = "Top 10 Browser", x = 'browser', y = "Number")
```
Most of the visitors used chrome browser. The most used browsers are chrome, safari, firefox. 

Here we explore the device category distribution. 

```{r}
dc = data.frame(table(train$deviceCategory))
dc  = dc%>% 
  mutate(percent = percent(Freq/sum(Freq)))

ggplot(dc,aes(x = reorder(Var1,-Freq) , y =Freq))+
  geom_bar(stat="identity",fill="#FF6666")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
          geom_text(aes(label=percent ), vjust=0) +
    labs(title = "Device Category", x = 'Device Category', y = "Number")
```

The most used device is desktop. 

```{r}
IM = data.frame(table(train$isMobile))
IM  = IM%>% 
  mutate(percent = percent(Freq/sum(Freq)))

ggplot(IM,aes(x = reorder(Var1,-Freq) , y =Freq))+
  geom_bar(stat="identity",fill="#FF6666")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
    geom_text(aes(label=percent ), vjust=0) +
    labs(title = "isMobile", x = 'isMobile', y = "Number")
```


Here we explore the device operating system. 

```{r}
train_os = train[which(train$operatingSystem%in%names(sort(table(train$operatingSystem),decreasing=TRUE)[1:10])),]
os = data.frame(table(train_os$operatingSystem))
os = os[-which(os$Freq==0),]
os  = os%>% 
  mutate(percent = percent(Freq/sum(Freq)))

ggplot(os,aes(x = reorder(Var1,-Freq) , y =Freq))+
  geom_bar(stat="identity",fill="#FF6666")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
      geom_text(aes(label=percent), vjust=0) +
    labs(title = "Top 10 Operating System", x = 'Operating System', y = "Number")
```
Most of the visitors used Mac. The most used operating systems are Mac, Windows, IOS, Android.  


Here we explore the device network domain. 

```{r}
train_nd = train[which(train$networkDomain%in%names(sort(table(train$networkDomain),decreasing=TRUE)[1:10])),]
networkDomain = data.frame(table(train_nd$networkDomain))
networkDomain = networkDomain[-which(networkDomain$Freq==0),]
networkDomain  = networkDomain%>% 
  mutate(percent = percent(Freq/sum(Freq)))

ggplot(networkDomain,aes(x = reorder(Var1,-Freq) , y =Freq))+
  geom_bar(stat="identity",fill="#FF6666")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
        geom_text(aes(label=percent), vjust=0) +
    labs(title = "Top 10 Network Domain", x = 'Network Domain', y = "Number")
```
The most used network domains are comcast, rr, verizon. 


Now we check these features correlation.

```{r}
ob = table(train_brow$browser,train_brow$operatingSystem)
ob = ob[,which(colSums(ob)!=0)]
ob = ob[which(rowSums(ob)!=0),]
ob = melt(ob)

ggplot(ob, aes(Var.1,Var.2)) + geom_tile(aes(fill = value), colour = "white")+
  scale_fill_gradient(low = "white",high = "steelblue")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
    geom_text(aes(label = round(value, 1))) +
    labs(title = "Browser vs. OperatingSystem ", x = 'Browser', y = "OperatingSystem")
```

The most frequent combiniation of the visits is using Chrome browser from Mac. 


```{r}
dd = table(train_brow$browser,train_brow$deviceCategory)
dd = dd[,which(colSums(dd)!=0)]
dd = dd[which(rowSums(dd)!=0),]
dd = melt(dd)

ggplot(dd, aes(Var.2,Var.1)) + geom_tile(aes(fill = value), colour = "white")+
  scale_fill_gradient(low = "white",high = "steelblue")+
      geom_text(aes(label = round(value, 1))) +
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
    labs(title = "Browser vs. Device Category ", x = 'Device Category', y = "Browser")

```

The most frequent combiniation of the visits is using Chrome browser on desktop. 

```{r}
mc = table(train_brow$isMobile,train_brow$deviceCategory)
mc = mc[,which(colSums(mc)!=0)]
mc = mc[which(rowSums(mc)!=0),]
mc = melt(mc)

ggplot(mc, aes(Var.1,Var.2)) + geom_tile(aes(fill = value), colour = "white")+
  scale_fill_gradient(low = "white",high = "steelblue")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
      geom_text(aes(label = round(value, 1))) +
    labs(title = "Device is Mobile vs. Device Category ", x = 'Device is Mobile ', y = "Device Category")
```

Now we explore geographical attributes of the visit. 

```{r}
city = data.frame(table(train$city))
city = city[city$Freq>mean(city$Freq),]
city = city%>% 
  mutate(percent = percent(Freq/sum(Freq)))


plot_ly(city,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = city$percent)%>%
  layout(title = "City",
         xaxis = list(title = "City"),
        yaxis = list(title = "Number"))
```

```{r}

reg = data.frame(table(train$region))
reg = reg[reg$Freq>mean(reg$Freq),]
reg = reg%>% 
  mutate(percent = percent(Freq/sum(Freq)))

# ggplot(reg,aes(x = reorder(Var1,-Freq) , y =Freq))+
#   geom_bar(stat="identity",fill="steelblue")+
#   theme(plot.title = element_text(hjust = 0.5),
#     axis.text.x = element_text(angle = 45, hjust = 1, size=10))+
#     labs(title = "Region", x = 'Region', y = "Number")

plot_ly(reg,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = reg$percent)%>%
  layout(title = "Region",
         xaxis = list(title = "Region"),
        yaxis = list(title = "Number"))

```


```{r}
metro = data.frame(table(train$metro))
metro = metro[metro$Freq>mean(metro$Freq),]
ggplot(metro,aes(x = reorder(Var1,-Freq) , y =Freq))+
  geom_bar(stat="identity",fill="steelblue")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=10))+
    labs(title = "Metro", x = 'metro', y = "Number")

```

