---
title: "Advanced Data Science-Final Project"
output:
  html_document:
    df_print: paged
---

The 80/20 rule has proven true for many businesses–only a small percentage of customers produce most of the revenue. As such, marketing teams are challenged to make appropriate investments in promotional strategies.

GStore
RStudio, the developer of free and open tools for R and enterprise-ready products for teams to scale and share work, has partnered with Google Cloud and Kaggle to demonstrate the business impact that thorough data analysis can have.

In this competition, you’re challenged to analyze a Google Merchandise Store (also known as GStore, where Google swag is sold) customer dataset to predict revenue per customer. Hopefully, the outcome will be more actionable operational changes and a better use of marketing budgets for those companies who choose to use data analysis on top of GA data

First we read the data into R
```{r}
library(tidyverse)
dat <- read_csv("/Users/chenhuan/Downloads/Data-3/train_US_1year_nojson.csv")
```
Then, take a glimpse at the data
```{r}
## Dimension of the dat
dim(dat)
## Take a glimpse at the dat
glimpse(dat)
## Summary of the dat
summary(dat)
## The head part of the dat
head(dat)

```
```{r}
library(ggplot2)
library(lubridate)
## Density of the transactionRevenue
dat %>% 
  select(fullVisitorId, totalTransactionRevenue) %>%
  group_by(fullVisitorId) %>% 
  summarise(total.revenue = sum(log(na.omit(totalTransactionRevenue)))) %>% 
  ggplot(aes(x = total.revenue) ) + 
  geom_density()

## Total number of visotor
visitor.num = length(unique(dat$fullVisitorId))
print(paste0("The total number of different visitors is ", visitor.num))

## The distribution of how browsers are used
dat %>% 
  group_by(browser) %>%
  summarise(n = n()) %>%
  top_n(10) %>%
  ggplot(aes(x = browser, y = n / sum(n))) +
  geom_bar(stat="identity") + 
  ggtitle("Distribution of the top 10 mostly used browsers") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  ylab("Percentage") + 
  xlab("Browser Type")

## The distribution of the device
dat %>% 
  group_by(deviceCategory) %>%
  summarise(n = n()) %>%
  top_n(10) %>%
  ggplot(aes(x = deviceCategory, y = n / sum(n))) +
  geom_bar(stat="identity") + 
  ggtitle("Distribution of the devices") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  ylab("Percentage") + 
  xlab("Browser Type")

## The distribution of the OS
dat %>% 
  group_by(operatingSystem) %>%
  summarise(n = n()) %>%
  top_n(10) %>%
  ggplot(aes(x = operatingSystem, y = n / sum(n))) +
  geom_bar(stat="identity") + 
  ggtitle("Distribution of the operating system") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  ylab("Percentage") + 
  xlab("Operating System")

## The trend of the revenue with respect to the time 
dat %>% 
  mutate(date = ymd(date)) %>%
  group_by(date) %>%
  summarise(total.revenue = sum(na.omit(totalTransactionRevenue))) %>%
  ggplot(aes(x = date, y = total.revenue)) + 
  geom_line(smooth = "line") + 
  ggtitle("Time Series of totalTransactionRevenue") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  ylab("totalTransactionRevenue") + 
  xlab("Data")

## Mobile or Unmobile
dat %>% 
  group_by(isMobile) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = isMobile, y = n / sum(n))) +
  geom_bar(stat="identity") + 
  ggtitle("Mobile or Unmobile") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  ylab("Percentage") + 
  xlab("IsMobile")




```

Study the geographical information
```{r}
library(usmap)

dat %>% 
  select(region) %>%
  group_by(region) %>%
  summarise(n = n())

  

plot_usmap(data = statepop, values = "pop_2015", lines = "red") + 
  scale_fill_continuous(name = "Population (2015)", label = scales::comma) + 
  theme(legend.position = "right")


```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

