---
title: "project"
author: "LUQIN GAN"
date: "12/3/2018"
output: html_document
---

## Data exploration

```{r, warning = FALSE, message = FALSE, tidy = TRUE}
library(ggplot2)
library(lubridate)
library(usmap)
library(tidyverse)
library(reshape)
library(knitr)
library(kableExtra)
library(plotly)
train = read.csv('/Users/chenhuan/Downloads/Data-3/train_US_1year_nojson.csv')

```

#### Check features with missing values 
```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
train$totalTransactionRevenue[which(is.na(train$totalTransactionRevenue))] = 0
miss = data.frame(Features = colnames(train), Number = sapply(1:ncol(train), function(x) length(which(is.na(train[,x])))), Percent = round(sapply(1:ncol(train), function(x) length(which(is.na(train[,x]))))/nrow(train),2))

miss_bar = miss[(which(miss$Percent!=0)),]
plot_ly(miss_bar,y=~reorder(Features,-Percent),x=~Percent,type = 'bar',text = miss_bar$Percent,orientation = 'h')%>%
  layout(title = "Missing data",
         xaxis = list(title = "Percent"),
        yaxis = list(title = "Features"))
# 
# plot_ly(miss_bar,x=~reorder(Features,-Percent),y=~Percent,type = 'bar',text = miss_bar$Percent)%>%
#   layout(title = "Missing data",
#          yaxis = list(title = "Percent"),
#          xaxis = list(title = "Features"))

miss %>%
  mutate_if(is.numeric, function(x) {
    cell_spec(x, bold = T, 
              color = spec_color(x, end = 0.9),
              font_size = spec_font_size(x))
  }) %>%
  kable(escape = F, align = "c",caption = "Missing data summary") %>%
  kable_styling(c("striped", "condensed"), full_width = F)

dat = train[,which(miss$Percent<0.8)]

# test_shiny = train[1:100,-c(32:35)]
# 
# write.csv(test_shiny,file='/Users/alice/Documents/yr2term1/data science/homework/google_revenue/test_shiny.csv')

```

We droped the features which have greater than 80% missing values. 

#### Channel Grouping 

Here we explore the marketing channel grouping distribution. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}

cg = data.frame(table(train$channelGrouping))
percent <- function(x, digits = 2, format = "f", ...) {
  paste0(formatC(100 * x, format = format, digits = digits, ...), "%")
}
cg  = cg%>% 
  mutate(percent = percent(Freq/sum(Freq)))

cg = cg[order(-cg$Freq),]

plot_ly(cg,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = cg$percent)%>%
  layout(title = "Channel Grouping",
         xaxis = list(title = "Channel Grouping"),
        yaxis = list(title = "Number"))
```
Over 40% of the visitors used organic search. 

Here we explore the browser distribution. 
```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}

names(sort(table(train$browser),decreasing=TRUE)[1:10])
train_brow = train[which(train$browser%in%names(sort(table(train$browser),decreasing=TRUE)[1:10])),]
brow = data.frame(table(train_brow$browser))
brow = brow[-which(brow$Freq==0),]
brow  = brow%>% 
  mutate(percent = percent(Freq/sum(Freq)))

plot_ly(brow,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = brow$percent)%>%
  layout(title = "Top 10 Browser",
         xaxis = list(title = "Browser"),
        yaxis = list(title = "Number"))
```
Most of the visitors used chrome browser. The most used browsers are chrome, safari, firefox. 

Here we explore the device category distribution. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
dc = data.frame(table(train$deviceCategory))
dc  = dc%>% 
  mutate(percent = percent(Freq/sum(Freq)))

plot_ly(dc,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = dc$percent)%>%
  layout(title = "Device Category",
         xaxis = list(title = "Device Category"),
        yaxis = list(title = "Number"))
colnames(train)
```

The most used device is desktop. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
IM = data.frame(table(train$isMobile))
IM  = IM%>% 
  mutate(percent = percent(Freq/sum(Freq)))

plot_ly(IM,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = IM$percent)%>%
  layout(title = "isMobile",
         xaxis = list(title = "isMobile"),
        yaxis = list(title = "Number"))

```


Here we explore the device operating system. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
train_os = train[which(train$operatingSystem%in%names(sort(table(train$operatingSystem),decreasing=TRUE)[1:10])),]
os = data.frame(table(train_os$operatingSystem))
os = os[-which(os$Freq==0),]
os  = os%>% 
  mutate(percent = percent(Freq/sum(Freq)))

plot_ly(os,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = os$percent)%>%
  layout(title = "Top 10 Operating System",
         xaxis = list(title = "Operating System"),
        yaxis = list(title = "Number"))

```
Most of the visitors used Mac. The most used operating systems are Mac, Windows, IOS, Android.  


Here we explore the device network domain. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
train_nd = train[which(train$networkDomain%in%names(sort(table(train$networkDomain),decreasing=TRUE)[1:10])),]
networkDomain = data.frame(table(train_nd$networkDomain))
networkDomain = networkDomain[-which(networkDomain$Freq==0),]
networkDomain  = networkDomain%>% 
  mutate(percent = percent(Freq/sum(Freq)))


plot_ly(networkDomain,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = networkDomain$percent)%>%
  layout(title = "Top 10 Network Domain",
         xaxis = list(title = "Network Domain"),
        yaxis = list(title = "Number"))
```
The most used network domains are comcast, rr, verizon. 

Now we check these features correlation.

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
ob = table(train$browser,train$operatingSystem)
ob = ob[,which(colSums(ob)!=0)]
ob = ob[which(rowSums(ob)!=0),]
ob = melt(ob)

ggplot(ob, aes(Var.1,Var.2)) + geom_tile(aes(fill = value), colour = "white")+
  scale_fill_gradient(low = "white",high = "steelblue")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
    geom_text(aes(label = round(value, 1))) +
    labs(title = "Browser vs. OperatingSystem ", x = 'Browser', y = "OperatingSystem")

```

The most frequent combiniation of the visits is using Chrome browser from Mac. 


```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
dd = table(train_brow$browser,train_brow$deviceCategory)
dd = dd[,which(colSums(dd)!=0)]
dd = dd[which(rowSums(dd)!=0),]
dd = melt(dd)

ggplot(dd, aes(Var.2,Var.1)) + geom_tile(aes(fill = value), colour = "white")+
  scale_fill_gradient(low = "white",high = "steelblue")+
      geom_text(aes(label = round(value, 1))) +
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
    labs(title = "Browser vs. Device Category ", x = 'Device Category', y = "Browser")

```

The most frequent combiniation of the visits is using Chrome browser on desktop. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
mc = table(train_brow$isMobile,train_brow$deviceCategory)
mc = mc[,which(colSums(mc)!=0)]
mc = mc[which(rowSums(mc)!=0),]
mc = melt(mc)

ggplot(mc, aes(Var.1,Var.2)) + geom_tile(aes(fill = value), colour = "white")+
  scale_fill_gradient(low = "white",high = "steelblue")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
      geom_text(aes(label = round(value, 1))) +
    labs(title = "Device is Mobile vs. Device Category ", x = 'Device is Mobile ', y = "Device Category")
```

Now we explore geographical attributes of the visit. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
city = data.frame(table(train$city))
city = city[city$Freq>mean(city$Freq),]
city = city%>% 
  mutate(percent = percent(Freq/sum(Freq)))


plot_ly(city,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = city$percent)%>%
  layout(title = "City",
         xaxis = list(title = "City"),
        yaxis = list(title = "Number"))
```

The most frequent city is Mountain View. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
reg = data.frame(table(train$region))
reg = reg[reg$Freq>mean(reg$Freq),]
reg = reg%>% 
  mutate(percent = percent(Freq/sum(Freq)))

plot_ly(reg,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = reg$percent)%>%
  layout(title = "Region",
         xaxis = list(title = "Region"),
        yaxis = list(title = "Number"))

```

The most frequent region is California. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
metro = data.frame(table(train$metro))
metro = metro[metro$Freq>mean(metro$Freq),]
metro = metro%>% 
  mutate(percent = percent(Freq/sum(Freq)))

plot_ly(metro,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = metro$percent)%>%
  layout(title = "Metro",
         xaxis = list(title = "Metro"),
        yaxis = list(title = "Number"))
```

The most frequent metros are San Francisco, New York, and Los Angeles. 


```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
# The revenue corresponds to the different states in USA
l <- list(color = toRGB("white"), width = 2)
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
map.visit <- dat %>%
  select(region,transactionRevenue) %>%
  group_by(region) %>%
  summarise(n = round(log(n()),3),rev = round(log(sum(as.numeric(na.omit(transactionRevenue)))+1),3)) %>%
  rename(full = region)


state.google <- statepop %>%
  select(fips, abbr, full) %>%
  left_join(map.visit, by = "full")

state.google$n[is.na(state.google$n)] = 0
# state.google$hover <- with(state.google, paste('State: ',full, '<br>', "Revenue: ", n))
state.google$hover <- with(state.google, paste(full))

plot_geo(state.google, locationmode = 'USA-states') %>%
  add_trace(
    z = ~n, text = ~hover, locations = ~abbr,
    color = ~n, colors = 'Greens'
  ) %>%
  colorbar(title = "log(Number of visit)") %>%
  layout(
    title = 'Visit per State (log scale)',
    geo = g
  )
```

The state with the largest number of visit is California, and the state with the second largest visit is New York. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}

plot_geo(state.google, locationmode = 'USA-states') %>%
  add_trace(
    z = ~rev, text = ~hover, locations = ~abbr,
    color = ~rev, colors = 'Blues'
  ) %>%
  colorbar(title = "log(TransactionRevenue) USD") %>%
  layout(
    title = 'Google transaction revenue by State (log scale)',
    geo = g
  )
```

Consistent with the number of visit, the state with the highest transaction revenue is California, and the state with the second highest transaction revenue is New York. 

Now we show the distribution of visit per state, considering only visits with non-zero transactions. 
```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
map.visit2 <- dat %>%
  select(region,transactionRevenue) %>%
  filter(transactionRevenue>0)%>%
  group_by(region) %>%
  summarise(n = round(log(n()),3),rev = round(log(sum(as.numeric(na.omit(transactionRevenue)))+1),3)) %>%
  rename(full = region)

state.google2 <- statepop %>%
  select(fips, abbr, full) %>%
  left_join(map.visit2, by = "full")

state.google2$n[is.na(state.google2$n)] = 0

state.google2$hover <- with(state.google2, paste(full))

plot_geo(state.google2, locationmode = 'USA-states') %>%
  add_trace(
    z = ~n, text = ~hover, locations = ~abbr,
    color = ~n, colors = 'Reds'
  ) %>%
  colorbar(title = "log(Number of visit) ") %>%
  layout(
    title = 'Visit per State in non-zero transactions (log scale)',
    geo = g
  )

```

The total numbers in each state are smaller, and the top 2 states with largest number of visits are California and New York. 

#### Date and time 
```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
dat = dat%>%
  mutate(date= ymd(date))

date = dat%>%
  select(date,transactionRevenue) %>%
  group_by((date)) %>%
  summarise(n = round(log(n()),3),rev = round(log(sum(as.numeric(na.omit(transactionRevenue)))+1),3))
colnames(date)[1] = 'Date'

date$hover <- with(date, paste('Date: ',Date, '<br>', "Visits: ", n,'<br>','Revenue: ', rev))

plot_ly(x = ~Date, y = ~rev, data = date, mode = 'lines', hoverinfo = 'text', text = date$hover)%>%
  layout(title = "Revenues per day (log scale)",
         xaxis = list(title = "Day"),
        yaxis = list(title = "Revenue"))

```

We plot the time series for transaction revenues. 


```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
year = dat%>%
  select(date,transactionRevenue) %>%
  group_by(year(date)) %>%
  summarise(n = round(log(n()),3),rev = round(log(sum(as.numeric(na.omit(transactionRevenue)))+1),3))

colnames(year)[1] = 'Year'
year$Year = as.character(year$Year)
ggplot(year,aes(x = Year , y = n))+
  geom_bar(stat="identity",fill="#FF6666")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
          geom_text(aes(label=n ), vjust=0) +
    labs(title = "Visit per year (log scale)", x = 'Year', y = "Number")

ggplot(year,aes(x = Year , y = rev))+
  geom_bar(stat="identity",fill="darkblue")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
          geom_text(aes(label=rev ), vjust=0) +
    labs(title = "Revenue per year (log scale)", x = 'Year', y = "Number")
```

For 2016 and 2017, the numbers of visit in each year are similar, as well as tansaction revenue. 


```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
mm <- c('January', 'February', 'March', 'April', 'May', 'June', 'July',
         'August', 'September', 'October', 'November', 'December')
month = dat%>%
  select(date,transactionRevenue) %>%
  group_by(month(date)) %>%
  summarise(n = round(log(n()),3),rev = round(log(sum(as.numeric(na.omit(transactionRevenue)))+1),3))
month = data.frame(Month = mm,month)

month$Month <- factor(month$Month, levels = month[['Month']])

month$hover <- with(month, paste('Month: ',Month, '<br>', "Visits: ", n,'<br>','Revenue: ', rev))

plot_ly(x = ~Month, y = ~rev, data = month,type = 'scatter', mode = 'lines', hoverinfo = 'text', text = month$hover)%>%
  layout(title = "Revenues per month (log scale)",
         xaxis = list(title = "Month"),
        yaxis = list(title = "Revenue"))
```

April has the highest revenue. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
week = dat%>%
  select(date,transactionRevenue) %>%
  group_by(weekdays(date)) %>%
  summarise(n = round(log(n()),3),rev = round(log(sum(as.numeric(na.omit(transactionRevenue)))+1),3))

colnames(week)[1] = 'Week'

# week$Week <- factor(week$Week, levels = week[['Week']])
# # 
# # 
week$Week <- factor(week$Week, levels = c("Monday","Tuesday","Wednesday", "Thursday","Friday","Saturday","Sunday"))

week$hover <- with(week, paste('Week: ',Week, '<br>', "Visits: ", n,'<br>','Revenue: ', rev))

order = c(2,6,7,5,1,3,4)
week = week[order,]

plot_ly(x = ~Week, y = ~rev, data = week,type = 'scatter', mode = 'lines', hoverinfo = 'text', text = week$hover)%>%
  layout(title = "Revenues per weekday (log scale)",
         xaxis = list(title = "Weekday"),
        yaxis = list(title = "Revenue"))
```

Revenue is higher in weekdays than weekends. 
