---
title: "project"
author: "LUQIN GAN"
date: "12/3/2018"
output: html_document
---

## Data exploration




```{r, warning = FALSE, message = FALSE, tidy = TRUE}

library(ggplot2)
library(lubridate)
library(usmap)
library(tidyverse)
library(reshape)
library(knitr)
library(kableExtra)
library(plotly)

#train = read.csv('/Users/alice/Documents/yr2term1/data science/project/Data/train_US_1year_nojson.csv')

train = read.csv('/Users/chenhuan/Downloads/Data-3/train_US_1year_nojson.csv')


```

#### Check features with missing values 



### change rev NA to  0 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}

train$totalTransactionRevenue[which(is.na(train$totalTransactionRevenue))] = 0

## count missing values
miss = data.frame(Features = colnames(train), Number = sapply(1:ncol(train), function(x) length(which(is.na(train[,x])))), Percent = round(sapply(1:ncol(train), function(x) length(which(is.na(train[,x]))))/nrow(train),2))


## delete no missing variables 
miss_bar = miss[(which(miss$Percent!=0)),]

### plot missing percent hist

miss_bar = miss[(which(miss$Percent!=0)),]

plot_ly(miss_bar,y=~reorder(Features,-Percent),x=~Percent,type = 'bar',text = miss_bar$Percent,orientation = 'h')%>%
  layout(title = "Missing data",
         xaxis = list(title = "Percent"),
        yaxis = list(title = "Features"))


##Missing data summary table 


miss %>%
  mutate_if(is.numeric, function(x) {
    cell_spec(x, bold = T, 
              color = spec_color(x, end = 0.9),
              font_size = spec_font_size(x))
  }) %>%
  kable(escape = F, align = "c",caption = "Missing data summary") %>%
  kable_styling(c("striped", "condensed"), full_width = F)

## delete var with > 80% missing 
dat = train[,which(miss$Percent<0.8)]

# test_shiny = train[1:100,-c(32:35)]
# 
# write.csv(test_shiny,file='/Users/alice/Documents/yr2term1/data science/homework/google_revenue/test_shiny.csv')

```

We droped the features which have greater than 80% missing values. 

#### Channel Grouping 

Here we explore the marketing channel grouping distribution. 


```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
## count channelGrouping


cg = data.frame(table(train$channelGrouping))

## percent function 
percent <- function(x, digits = 2, format = "f", ...) {
  paste0(formatC(100 * x, format = format, digits = digits, ...), "%")
}

## calculate percent 
cg  = cg%>% 
  mutate(percent = percent(Freq/sum(Freq)))
# rank by freq 
cg = cg[order(-cg$Freq),]

cg = cg[order(-cg$Freq),]

plot_ly(cg,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = cg$percent)%>%
  layout(title = "Channel Grouping",
         xaxis = list(title = "Channel Grouping"),
        yaxis = list(title = "Number"))
```
Over 40% of the visitors used organic search. 

Here we explore the browser distribution. 


## count browser


```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
# keep top 10 browser 

names(sort(table(train$browser),decreasing=TRUE)[1:10])

train_brow = train[which(train$browser%in%names(sort(table(train$browser),decreasing=TRUE)[1:10])),]
brow = data.frame(table(train_brow$browser))

# delete freq = 0 rows 
brow = brow[-which(brow$Freq==0),]
## calculate percent 

## calculate percent 
brow  = brow%>% 
  mutate(percent = percent(Freq/sum(Freq)))

plot_ly(brow,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = brow$percent)%>%
  layout(title = "Top 10 Browser",
         xaxis = list(title = "Browser"),
        yaxis = list(title = "Number"))
```
Most of the visitors used chrome browser. The most used browsers are chrome, safari, firefox. 

Here we explore the device category distribution. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
dc = data.frame(table(train$deviceCategory))
## calculate percent 
dc  = dc%>% 
  mutate(percent = percent(Freq/sum(Freq)))

plot_ly(dc,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = dc$percent)%>%
  layout(title = "Device Category",
         xaxis = list(title = "Device Category"),
        yaxis = list(title = "Number"))
colnames(train)
```

The most used device is desktop. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
IM = data.frame(table(train$isMobile))
                
## calculate percent 
IM  = IM%>% 
  mutate(percent = percent(Freq/sum(Freq)))

plot_ly(IM,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = IM$percent)%>%
  layout(title = "isMobile",
         xaxis = list(title = "isMobile"),
        yaxis = list(title = "Number"))

```


Here we explore the device operating system. 



# explore the device operating system

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}

train_os = train[which(train$operatingSystem%in%names(sort(table(train$operatingSystem),decreasing=TRUE)[1:10])),]
os = data.frame(table(train_os$operatingSystem))

# delete 0 rows  
os = os[-which(os$Freq==0),]
## calculate percent 
os  = os%>% 
  mutate(percent = percent(Freq/sum(Freq)))

plot_ly(os,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = os$percent)%>%
  layout(title = "Top 10 Operating System",
         xaxis = list(title = "Operating System"),
        yaxis = list(title = "Number"))

```
Most of the visitors used Mac. The most used operating systems are Mac, Windows, IOS, Android.  


Here we explore the device network domain. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
train_nd = train[which(train$networkDomain%in%names(sort(table(train$networkDomain),decreasing=TRUE)[1:10])),]
networkDomain = data.frame(table(train_nd$networkDomain))
networkDomain = networkDomain[-which(networkDomain$Freq==0),]
## calculate percent 
networkDomain  = networkDomain%>% 
  mutate(percent = percent(Freq/sum(Freq)))


plot_ly(networkDomain,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = networkDomain$percent)%>%
  layout(title = "Top 10 Network Domain",
         xaxis = list(title = "Network Domain"),
        yaxis = list(title = "Number"))
```
The most used network domains are comcast, rr, verizon. 

Now we check these features correlation.


```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
##  browser and operatingSystem correlation.
ob = table(train$browser,train$operatingSystem)

## delete 0 columns
ob = ob[,which(colSums(ob)!=0)]
## delete 0 rows
ob = ob[which(rowSums(ob)!=0),]

## change to long format
ob = melt(ob)

ggplot(ob, aes(Var.1,Var.2)) + geom_tile(aes(fill = value), colour = "white")+
  scale_fill_gradient(low = "white",high = "steelblue")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
    geom_text(aes(label = round(value, 1))) +
    labs(title = "Browser vs. OperatingSystem ", x = 'Browser', y = "OperatingSystem")

```

The most frequent combiniation of the visits is using Chrome browser from Mac. 


```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
dd = table(train_brow$browser,train_brow$deviceCategory)

## delete 0 columns
dd = dd[,which(colSums(dd)!=0)]
## delete 0 rows
dd = dd[which(rowSums(dd)!=0),]
## change to long format
dd = melt(dd)

ggplot(dd, aes(Var.2,Var.1)) + geom_tile(aes(fill = value), colour = "white")+
  scale_fill_gradient(low = "white",high = "steelblue")+
      geom_text(aes(label = round(value, 1))) +
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
    labs(title = "Browser vs. Device Category ", x = 'Device Category', y = "Browser")

```

The most frequent combiniation of the visits is using Chrome browser on desktop. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
mc = table(train_brow$isMobile,train_brow$deviceCategory)
## delete 0 columns
mc = mc[,which(colSums(mc)!=0)]
## delete 0 rows
mc = mc[which(rowSums(mc)!=0),]
## change to long format
mc = melt(mc)

ggplot(mc, aes(Var.1,Var.2)) + geom_tile(aes(fill = value), colour = "white")+
  scale_fill_gradient(low = "white",high = "steelblue")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
      geom_text(aes(label = round(value, 1))) +
    labs(title = "Device is Mobile vs. Device Category ", x = 'Device is Mobile ', y = "Device Category")
```

Now we explore geographical attributes of the visit. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
city = data.frame(table(train$city))

## keep cities with have greater than average visits 
city = city[city$Freq>mean(city$Freq),]

## calculate percent 
city = city%>% 
  mutate(percent = percent(Freq/sum(Freq)))


plot_ly(city,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = city$percent)%>%
  layout(title = "City",
         xaxis = list(title = "City"),
        yaxis = list(title = "Number"))
```

The most frequent city is Mountain View. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
reg = data.frame(table(train$region))

## keep regions with have greater than average visits 
reg = reg[reg$Freq>mean(reg$Freq),]
## calculate percent 
reg = reg%>% 
  mutate(percent = percent(Freq/sum(Freq)))

plot_ly(reg,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = reg$percent)%>%
  layout(title = "Region",
         xaxis = list(title = "Region"),
        yaxis = list(title = "Number"))

```

The most frequent region is California. 

```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
metro = data.frame(table(train$metro))
## keep metros with have greater than average visits 
metro = metro[metro$Freq>mean(metro$Freq),]
## calculate percent 
metro = metro%>% 
  mutate(percent = percent(Freq/sum(Freq)))

plot_ly(metro,x=~reorder(Var1,-Freq),y=~Freq,type = 'bar',text = metro$percent)%>%
  layout(title = "Metro",
         xaxis = list(title = "Metro"),
        yaxis = list(title = "Number"))
```

The most frequent metros are San Francisco, New York, and Los Angeles. 


```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}

# The revenue corresponds to the different states in USA
l <- list(color = toRGB("white"), width = 2)
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)


## count rev and visit in usa 


map.visit <- dat %>%
  select(region,transactionRevenue) %>%
  group_by(region) %>%
  summarise(n = round(log(n()),3),rev = round(log(sum(as.numeric(na.omit(transactionRevenue)))+1),3)) %>%
  rename(full = region)

## join maps table by full 
state.google <- statepop %>%
  select(fips, abbr, full) %>%
  left_join(map.visit, by = "full")

## change na to 0
state.google$n[is.na(state.google$n)] = 0
# state.google$hover <- with(state.google, paste('State: ',full, '<br>', "Revenue: ", n))
state.google$hover <- with(state.google, paste(full))

plot_geo(state.google, locationmode = 'USA-states') %>%
  add_trace(
    z = ~n, text = ~hover, locations = ~abbr,
    color = ~n, colors = 'Greens'
  ) %>%
  colorbar(title = "log(Number of visit)") %>%
  layout(
    title = 'Visit per State (log scale)',
    geo = g
  )
```

The state with the largest number of visit is California, and the state with the second largest visit is New York. 




```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}

## plot rev by state in map 

plot_geo(state.google, locationmode = 'USA-states') %>%
  add_trace(
    z = ~rev, text = ~hover, locations = ~abbr,
    color = ~rev, colors = 'Blues'
  ) %>%
  colorbar(title = "log(TransactionRevenue) USD") %>%
  layout(
    title = 'Google transaction revenue by State (log scale)',
    geo = g
  )
```

Consistent with the number of visit, the state with the highest transaction revenue is California, and the state with the second highest transaction revenue is New York. 

Now we show the distribution of visit per state, considering only visits with non-zero transactions. 


```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
#considering only visits with non-zero transaction 

map.visit2 <- dat %>%
  select(region,transactionRevenue) %>%
  filter(transactionRevenue>0)%>%
  group_by(region) %>%
  summarise(n = round(log(n()),3),rev = round(log(sum(as.numeric(na.omit(transactionRevenue)))+1),3)) %>%
  rename(full = region)

### join map data by full
state.google2 <- statepop %>%
  select(fips, abbr, full) %>%
  left_join(map.visit2, by = "full")

## delete revenue = 0 visits 
state.google2$n[is.na(state.google2$n)] = 0

## add hover info 
state.google2$hover <- with(state.google2, paste(full))

### plot visit per state with non-zero rev 
plot_geo(state.google2, locationmode = 'USA-states') %>%
  add_trace(
    z = ~n, text = ~hover, locations = ~abbr,
    color = ~n, colors = 'Reds'
  ) %>%
  colorbar(title = "log(Number of visit) ") %>%
  layout(
    title = 'Visit per State in non-zero transactions (log scale)',
    geo = g
  )

```

The total numbers in each state are smaller, and the top 2 states with largest number of visits are California and New York. 

#### Date and time 
```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
##### change format to data 

dat = dat%>%
  mutate(date= ymd(date))
## summarize revenue per day 
date = dat%>%
  select(date,transactionRevenue) %>%
  group_by((date)) %>%
  summarise(n = round(log(n()),3),rev = round(log(sum(as.numeric(na.omit(transactionRevenue)))+1),3))
colnames(date)[1] = 'Date'
## add hover information 
date$hover <- with(date, paste('Date: ',Date, '<br>', "Visits: ", n,'<br>','Revenue: ', rev))

plot_ly(x = ~Date, y = ~rev, data = date, mode = 'lines', hoverinfo = 'text', text = date$hover)%>%
  layout(title = "Revenues per day (log scale)",
         xaxis = list(title = "Day"),
        yaxis = list(title = "Revenue"))

```

We plot the time series for transaction revenues. 


```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
## summarize revenue per year (2016,2017)

year = dat%>%
  select(date,transactionRevenue) %>%
  group_by(year(date)) %>%
  summarise(n = round(log(n()),3),rev = round(log(sum(as.numeric(na.omit(transactionRevenue)))+1),3))

colnames(year)[1] = 'Year'
year$Year = as.character(year$Year)
## plot visit number per year
ggplot(year,aes(x = Year , y = n))+
  geom_bar(stat="identity",fill="#FF6666")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
          geom_text(aes(label=n ), vjust=0) +
    labs(title = "Visit per year (log scale)", x = 'Year', y = "Number")

## plot revenue per year
ggplot(year,aes(x = Year , y = rev))+
  geom_bar(stat="identity",fill="darkblue")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size=13))+
          geom_text(aes(label=rev ), vjust=0) +
    labs(title = "Revenue per year (log scale)", x = 'Year', y = "Number")
```

For 2016 and 2017, the numbers of visit in each year are similar, as well as tansaction revenue. 


```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
mm <- c('January', 'February', 'March', 'April', 'May', 'June', 'July',
         'August', 'September', 'October', 'November', 'December')
###############  summarize transaction revenue per month 
month = dat%>%
  select(date,transactionRevenue) %>%
  group_by(month(date)) %>%
  summarise(n = round(log(n()),3),rev = round(log(sum(as.numeric(na.omit(transactionRevenue)))+1),3))
month = data.frame(Month = mm,month)

month$Month <- factor(month$Month, levels = month[['Month']])
###### add hover information to plotly
month$hover <- with(month, paste('Month: ',Month, '<br>', "Visits: ", n,'<br>','Revenue: ', rev))

plot_ly(x = ~Month, y = ~rev, data = month,type = 'scatter', mode = 'lines', hoverinfo = 'text', text = month$hover)%>%
  layout(title = "Revenues per month (log scale)",
         xaxis = list(title = "Month"),
        yaxis = list(title = "Revenue"))
```

April has the highest revenue. 


```{r, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE}
###############  summarize transaction revenue per weekday 

week = dat%>%
  select(date,transactionRevenue) %>%
  group_by(weekdays(date)) %>%
  summarise(n = round(log(n()),3),rev = round(log(sum(as.numeric(na.omit(transactionRevenue)))+1),3))

colnames(week)[1] = 'Week'

week$Week <- factor(week$Week, levels = c("Monday","Tuesday","Wednesday", "Thursday","Friday","Saturday","Sunday"))

###### add hover information to plotly
week$hover <- with(week, paste('Week: ',Week, '<br>', "Visits: ", n,'<br>','Revenue: ', rev))

######## order by weekdays 
order = c(2,6,7,5,1,3,4)
week = week[order,]

plot_ly(x = ~Week, y = ~rev, data = week,type = 'scatter', mode = 'lines', hoverinfo = 'text', text = week$hover)%>%
  layout(title = "Revenues per weekday (log scale)",
         xaxis = list(title = "Weekday"),
        yaxis = list(title = "Revenue"))
```

Revenue is higher in weekdays than weekends. 
